\documentclass[11pt]{article}

\usepackage{epsfig, graphics, graphicx}
\usepackage{latexsym}
\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{amsmath,amssymb,enumerate,comment}

\newcommand{\qed}{\quad \ensuremath{\blacksquare}}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\prox}{\operatorname{prox}}
\newcommand{\conv}{\operatorname{conv}}
\newcommand{\op}{{\operatorname{op}}}

\begin{document}
\twocolumn
\section{Mastery set [25 points]\\(Aaditya)}
\paragraph{Lemma 1:} If $f : D \to \R$ is convex and $m \in \R$, then
$U_f(m) := \{ x \in D : f(x) \leq m \}$ is convex.

\paragraph{Proof:} For $x,y \in U_f(m), \theta \in [0,1]$, if
$z=\theta x+(1-\theta)y$, then
\[f(z) \leq \theta f(x) + (1 - \theta)f(y) \leq m,\] so
$z \in U_f(m)$. \qed

\paragraph{A [2 + 2]} If $S = \emptyset$, then $S$ is trivially convex.
Otherwise, for $x \in S$, $S = U_f(f(x))$ is convex, by Lemma 1. If $f$ is
strictly convex and $\exists x,y \in S$ with $x \neq y$, for
$\theta \in (0,1)$, $z = \theta x + (1 - \theta)y$,
\[f(z) < \theta f(x) + (1 - \theta)f(y) = f(x),\]
contradicting the fact that $x$ minimizes $f$.

\paragraph{B [2 + 2]} If $A$ is orthogonal, $A^TA = I$, whose only singular
value is $1$. For
\[A =
\begin{bmatrix}
    1 & 0 \\
    0 & 1
\end{bmatrix},
B =
\begin{bmatrix}
    0 & 1 \\
    1 & 0
\end{bmatrix},
\]
both $A$ and $B$ are orthogonal, but \fbox{$\tfrac12A + \tfrac12B$} is not
orthogonal.

\paragraph{C [2 + 2 + 2]} By Lemma 1, the unit ball $U_{\|\cdot\|}(1)$ is
convex. Since, by part B, all orthogonal matrices are in the unit ball, it
follows that any convex combination of orthogonal matrices is in the unit ball.
By the triangle inequality, for any norm $\|\cdot\|$,
\begin{align*}
                \|x\| & = \|x - y + y\| \leq \|x - y\| + \|y\|  \\
\mbox{ and }    \|y\| & = \|y - x + x\| \leq \|x - y\| + \|x\|,
\end{align*}
which implies
\[-\|x - y\| \leq \|x\| - \|y\| \leq \|x - y\|,\]
and so $|\|x\| - \|y\|| \leq \|x - y\|$ (this statement is precisely the
Reverse Triangle Inequality. Hence, any norm is Lipschitz with respect to
itself, with Lipschitz constant \fbox{$L = 1$.} \qed

\paragraph{D [5]} We already showed that the convex hull of orthogonal matrices
in contained in the unit ball of $\|\cdot\|_\op$.

If $A$ satisfies
$\|A\|_\op \leq 1$, then $A$ has a singular value decomposition
$A = U\Sigma V^T$, where $U$ and $V$ are orthogonal and $\Sigma$ is diagonal
and has entries in $[-1,1]$. Let $\D$ denote the set of diagonal matrices with
diagonal entries in $\{-1,1\}$. It is geometrically obvious that
$\conv\left(\{-1,1\}^n\right)~=~[-1,1]^n$, and hence $\Sigma \in \conv(\D)$.
Also, any $D \in \D$ is clearly orthogonal. Thus, since
$\Sigma = \sum_{D \in \D} \theta_D D$, where $\sum_{D \in \D} \theta_D = 1$,
\[A
    = U \left( \sum_{D \in \D} \theta_D D \right) V^T
    = \sum_{D \in \D} \theta_D UDV^T
\]
is in the convex hull of orthogonal matrices (as any product of orthogonal
matrices is itself orthogonal). \qed

\paragraph{E [4+2]} Rearranging the first-order definition of strong convexity,
\begin{align*}
f(y) - \langle \nabla f(x), y - x \rangle
    & \geq f(x) + \frac{\lambda}{2}\|y - x\|_2^2    \\
f(x) + \langle \nabla f(y), y - x \rangle
    & \geq f(y) + \frac{\lambda}{2}\|y - x\|_2^2.
\end{align*}
Adding these inequalities and cancelling terms,
\[\langle \nabla f(y) - \nabla f(x), y - x \rangle
    \geq \lambda\|y - x\|_2^2.\]
Then, by Cauchy-Schwarz,
\begin{align*}
\|\nabla f(y) - \nabla f(x)\|
 &  \geq \frac{\langle \nabla f(y) - \nabla f(x), y - x \rangle}{\|y - x\|_2}\\
 &  \geq + \lambda\|y - x\|_2. \qed
\end{align*}
If $\nabla f$ is Lipschitz with Lipschitz constant $L$, then
\[L\|y - x\|_2 \geq \|\nabla f(y) - \nabla f(x)\| \geq + \lambda\|y - x\|_2,\]
so \fbox{$L \geq \lambda$.}
\end{document}
